Constrained Optimization-Based Neuro-Adaptive Control (CoNAC) for Uncertain Euler-Lagrange Systems Under Weight and Input Constraints

### 2025-03-16 manuscript.tex
modifying

\newcommand*{\sat}{\mathcal{S}}
\newcommand*{\fe}{\mv r}
\newcommand*{\ptfrac}[2]{\tfrac{\partial #1}{\partial #2}}
\newcommand*{\dtfrac}[2]{\tfrac{\der #1}{\der #2}}
\newcommand*{\q}{\mv q}
\newcommand*{\dq}{\dot{\mv q}}
\newcommand*{\ddq}{\ddot{\mv q}}
\newcommand*{\qd}{\mv q_d}
\newcommand*{\dqd}{\dot{\mv q_d}}
\newcommand*{\ddqd}{\ddot{\mv q_d}}

\newcommand*{\NN}{\mv\Phi}
\newcommand*{\act}{\mv\phi}
\newcommand*{\wth}{\mv\theta}
\newcommand*{\wV}{\mm V}

@@

@@

\IEEEPARstart{M}{any} engineering systems, including those in aerospace, robotics, and automotive applications, can be modeled using Euler-Lagrange systems. These systems are governed by dynamic equations derived from energy principles and describe the motion of mechanical systems with constraints. In practice, however, such systems often exhibit uncertainties due to unmodeled dynamics, parameter variations, or external disturbances. These uncertainties can significantly degrade control performance and, in some cases, lead to instability. To address these challenges, adaptive control methods have been widely employed to ensure robust performance in the presence of system uncertainties \cite{RN4}, \cite{RN2}.

More recently, neuro-adaptive control approaches have been introduced to approximate unknown system dynamics or entire control laws using neural networks (NNs) \cite{RN1}. NNs are well-known for their universal approximation property, which allows them to approximate any smooth function over a compact set with minimal error. Various types of NNs have been utilized in neuro-adaptive control, including simpler architectures like single-hidden layer (SHL) neural networks \cite{RN29}, \cite{RN44} and radial basis function (RBF) neural networks \cite{RN26}, \cite{RN10}, as well as more complex models like deep neural networks (DNNs) \cite{RN16} and their variations. SHL and RBF NNs are often employed to approximate uncertain system dynamics or controllers due to their simplicity \cite{RN44, RN56, RN3, RN41}, while DNNs offer greater expressive power, making them more effective for complex system approximations \cite{RN25}. Additionally, variations of DNNs, such as long short-term memory (LSTM) networks for time-varying dynamics \cite{RN14} and physics-informed neural networks (PINNs) for leveraging physical system knowledge \cite{RN15}, have further extended the capabilities of neuro-adaptive control systems.

@@

However, two significant challenges persist in using NNs for adaptive control. First, the boundedness of NN weights is not inherently guaranteed, which can result in unbounded outputs. When NN outputs are used directly in the control law, this may lead to excessive control inputs, violating input constraints. Such constraints are commonly encountered in industrial systems, where actuators are limited by physical and safety requirements in terms of amplitude, rate, or energy \cite{RN18}. Failing to address these constraints can degrade control performance or even destabilize the system.

Therefore, addressing these two key issues—ensuring weight boundedness and satisfying input constraints—is essential for the reliable design of neuro-adaptive controllers. The following section will provide a detailed review of the existing solutions to these challenges.

@@

A common challenge in neuro-adaptive control is maintaining the boundedness of the NN weights to ensure stability. In many studies, projection operators were employed to enforce upper bounds on the weight norms, ensuring that the weights do not grow unboundedly. For example, in \cite{RN16, RN14, RN11}, projection operators were used to constrain the weight norms to remain below predetermined constants. However, these constants were often selected as large as possible due to the lack of theoretical guarantees regarding the global optimality of the weight values. While this approach ensured that the NN remained stable, it did not necessarily result in optimal performance.

In addition to projection operators, some studies utilized modification techniques like $\sigma$-modification \cite{RN10} and $\epsilon$-modification \cite{RN41, RN3}. These methods ensured that the NN weights remained within an invariant set by incorporating stabilizing functions into the adaptation law. Although these techniques were effective in ensuring boundedness and avoiding weight divergence, they similarly lacked a formal analysis of the optimality of the adapted weights, leaving room for improvement in terms of performance optimization.

@@

To address input saturation, many studies introduced auxiliary systems. These systems mitigated the effects of control input saturation by modifying the control strategy when saturation occurred. For instance, in \cite{RN55, RN56, RN3}, auxiliary states were generated whenever input saturation was detected, and these states were incorporated into the adaptation law to adjust the NN weights accordingly. This approach helped the controller reduce input saturation by indirectly regulating the auxiliary states.

Alternatively, auxiliary states can also be used as feedback terms in the control law to directly compensate for the effects of input saturation constraints, as demonstrated in \cite{RN34, RN38, RN37}. In \cite{RN41}, the NN was used to approximate the desired control input, compensating for input saturation. However, these approaches typically handle input bound constraints on a per-input basis (\ie applying bounds to each scalar control variable individually), and may not account for more complex, nonlinear constraints, like input norm constraints, which are commonly found in physical systems such as robotic actuators or motor systems.

@@

To overcome these limitations, constrained optimization offers a promising approach. By formulating the neuro-adaptive control problem as an optimization problem with constraints, it is possible to adapt the NN weights while minimizing an objective function (e.g., tracking error) subject to both weight norm and input constraints. Constrained optimization provides a theoretical framework for defining optimality and presents numerical methods for finding solutions that satisfy the constraints \cite{RN22}.

In existing literature, constrained optimization techniques, such as the Augmented Lagrangian Method (ALM) \cite{RN60} and the Alternating Direction Method of Multipliers (ADMM) \cite{RN58, RN59}, have been used to train NNs offline. These methods impose constraints to address issues like gradient vanishing in backpropagation. However, to the best of the authors' knowledge, no prior work has applied constrained optimization to adaptive control systems with real-time weight adaptation. This gap suggests that constrained optimization could be key to addressing both weight norm boundedness and input constraints in a unified, theoretically grounded framework, particularly in real-time neuro-adaptive control.

@@

The main contributions of this study are listed as follows:
\begin{itemize}
    \item A constrained optimization-based neuro-adaptive controller (CoNAC) is developed using a DNN, where input constraints and the boundedness of NN weights are formulated as inequality constraints within the adaptation process.
    \item The weight adaptation laws in CoNAC are derived using constrained optimization theory to minimize the objective function while satisfying the inequality constraints. The adaptation laws ensure convergence of the weights to the first-order optimality conditions, specifically the Karush-Kuhn-Tucker (KKT) conditions.
    \item The forward sensitivity method is applied in CoNAC to accurately calculate the gradient of the objective function.
    %, enabling more precise real-time adaptation.
\end{itemize}

@@

The remainder of this paper is organized as follows. 
Section \ref{sec: Problem Formulation} presents the target system and control objective.
Section \ref{sec:ctrl design} introduces the proposed controller and the architecture of DNN in the controller. 
In Section \ref{sec:adap_laws}, the weight adaptation laws are developed.
Candidates of the weight and input constraints are presented in Section \ref{sec:cstr candidates}.
Section \ref{sec:stability} analyzes the stability of the proposed controller.
A comparative study of the four selected controllers, including the proposed controller, is reported in Section \ref{sec:sim}.
Finally, Section \ref{sec:conclusion} concludes the paper and discusses potential future work.
% The appendices provide the proof of the lemma used in the stability analysis.

%  SECTION PROBLEM FORMULATION ============================
\section{Problem Formulation}\label{sec: Problem Formulation}

\subsection{Notation}
In this study, the following notation is used:

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite{RN17}.
    \item $\mv x=[x_i]_{i\in[1,\cdots,n]}\in\R^n$ and $\mm A\triangleq [a_{ij}]_{i\in[1,\cdots,n],j\in[1,\cdots ,]}\in\R^{n\times m}$ denotes a vector and a matrix.
    % \item $x_{(i)}$ denotes the $i\textsuperscript{th}$ element of the vector $x\in\mathbb R^n$.    
    % \item $A_{(i,j)}$ is the $i\textsuperscript{th}$ row-$j\textsuperscript{th}$ column element of matrix $A$.
    \item $\text{row}_i(A)$ denotes the $i\textsuperscript{th}$ row of the matrix $A\in\R^{n\times m}$. 
    % \item $\text{col}_i(A)$ and $\text{row}_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\R^{n\times m}$, respectively.
    \item $\textit{vec}(A)\triangleq [\text{row}_1(A^\top)  ,\cdots,\text{row}_m(A^\top)  ]^\top   $ for $A\in\R^{n\times m}$.
    % \item $\textit{vec}(A)\triangleq [\text{col}_1(A)^\top  ,\cdots,\text{col}_m(A)^\top  ]^\top   $ for $A\in\R^{n\times m}$.
    \item $\lambda_\text{min}(A)$ denotes the minimum eigenvalue of the matrix $A\in\R^{n\times n}$.
\end{itemize}

@@

Consider an uncertain Euler-Lagrange system modeled as
\begin{equation}
    \mm M\ddq + \mm V_m + \mm F + \mm G + \mv\tau_d
    =
    \sat(\mv\tau)
    \label{eq:sys1}
\end{equation}
where $\q\in \R^n$ denotes the generalized coordinate, $\mv \tau_d$ represents disturbance and $\mv\tau\in\R^n$ denotes the control input. 
The terms $\mm M\triangleq\mm M(\q)\in\R^{n\times n}$, $\mm V_m\triangleq\mm V_m(\q,\dq)\in\R^{n\times n}$, $\mm F\triangleq\mm F(\dq)\in\R^{n}$ and $\mm G\triangleq\mm G(\q)$ represent the unknown inertia matrix, Coriolis/centripetal matrix, friction terms and gravity vector
The function $\sat(\cdot):\R^n\to\R^n$ represents the inherent physical limitations of the actuators and is defined as 
\begin{equation}
    \sat(\mv\tau)
    \triangleq
    \begin{cases}
        \mv\tau, 
        & 
        \text{if } \Vert\mv\tau\Vert\le \bar\tau,
        \\
        \bar\tau
        \cdot
        \tfrac{\mv\tau}{\Vert\mv\tau\Vert}, 
        & 
        \text{otherwise},
    \end{cases}
\end{equation}
where $\bar\tau$ denotes the maximum norm of control input.
To account for these limitations, it is essential to incorporate physically motivated constraints into the controller design.
Section \ref{sec:cstr candidates} introduces candidate constraints that can be applied to ensure compliance with these physical limitations.

@@

\begin{prop} \cite{Lewis:1998aa}
    The Coriolis/centripetal matrix $\mm V_m$ can always be selected so that the matrix $\dot{\mm M}-2\mm V_m$ is skew-symmetric.
    \label{prop:skew}
\end{prop}

\begin{prop}
    The disturbance $\mv\tau_d$ is bounded so that $\norm{\mv\tau_d}\le \bar\tau_d$.
\end{prop}

@@

The architecture of the proposed CoNAC is illustrated in Fig.~\ref{fig: controller}, consisting of: a DNN that functions as a neuro-adaptive controller, and a weight optimizer for the DNN. 
Section \ref{sec:NAC} presents the neuro-adaptive controller along with the reference generator, and Section \ref{NN definition} defines the DNN model. The weight optimizer will be detailed in Section \ref{sec:adap_laws}.

\begin{figure*}[!t]
    \centering
    \includegraphics[width=0.7\linewidth]{fig/Controller.drawio.png}
    % \includesvg[width=0.75\linewidth]{Controller.drawio.svg}
    \caption{Architecture of the constrained optimization-based neuro-adaptive controller (CoNAC).}
    \label{fig: controller}
\end{figure*}

\subsection{Neuro-Adaptive Controller Design}\label{sec:NAC}

First, let us define the filtered tracking error $\fe\in\R^n$ as 
\begin{equation}
    \fe \triangleq \dot{\mv e}-\mm\Lambda \mv e
    ,
    \label{eq:err:filtered}
\end{equation}
where $\mv e \triangleq \q - \qd$ denotes the tracking error, $\dot{\mv e} \triangleq \dq - \dqd$ represents the derivative of the tracking error, and $\mm\Lambda\in\R^{n\times n}$ is a user-designed positive definite matrix.
Since \eqref{eq:err:filtered} is stable system, it implies that $\mv e$ is bounded if $\fe$ is bounded.

Using $\fe$, the system dynamics \eqref{eq:sys1} can be rewritten as
\begin{equation}
    \mm M \dot\fe
    =
    -\mm V_m \fe
    -\mm K \fe
    + \mv f
    -\mv \tau_d + \sat(\mv\tau)
    ,
\end{equation}
where $\mm K\in\R^{n\times n}$ denotes arbitrary symmetric positive definite matrix and $
    \mv f\triangleq \mv f(\q,\dq,\dqd,\ddqd)
    =
    \mm K \fe
    +\mm M(-\ddqd+\mm\Lambda\dot{\mv e})
    +
    \mm V_m(-\dqd+\mm\Lambda\mv e)
    -
    \mm F
    -
    \mm G
    \in\R^n
$ denotes the lumped system uncertainty.

Consider the Lyapunov function $\mathcal V_1\triangleq \tfrac{1}{2}\fe^\top \mm M\fe$. 
Invoking Property \ref{prop:skew}, the time derivative of $\mathcal V_1$ is
\begin{equation}
    \begin{aligned}
        \dot{\mathcal V}_1
        =&
        \fe^\top \mm M (
            -\mm V_m \fe -\mm K \fe + \mv f
            -\mv \tau_d + \sat(\mv\tau)
        )
        +
        \tfrac{1}{2}
        \fe^\top \dot{\mm M}\fe
        \\
        =&
        -
        \fe^\top \mm K \fe 
        +
        \fe^\top (
            \mv f+\sat(\mv\tau)
            -\mv \tau_d
        )
        +
        \tfrac{1}{2}
        \fe^\top(\dot{\mm M}- 2\mm V_m)\fe
        \\
        \le&
        -\lambda_\text{min}(\mm K)\norm{\fe}^2
        +
        \bar\tau_d\norm{\fe}
        +
        \fe^\top (\mv f+\sat(\mv \tau))
        \\
        \le&
        -\lambda_\text{min}(\mm K)\norm{\fe}^2
        +
        \bar\tau_d\norm{\fe}
        +
        \fe^\top (\sat(\mv\tau)-\mv\tau^*)
        ,
    \end{aligned}
    \label{eq:lya1:dot1}
\end{equation}
where $\mv\tau^*\triangleq -\mv f$ is the ideal control input. 
Therefore, one can conclude that $\fe$ is exponentially stable so that $\lim_{t\to\infty}\norm{\fe}=\tfrac{\bar\tau_d}{\lambda_\text{min}(\mm K)}$, if $\mv \tau=-\mv\tau^*$ can be realized.
However, $\mv \tau^*$ is not available in practice, since $\mv f$ is unknown.

To overcome this issue, a DNN is employed to approximate $\mv\tau^*$.
Let $\NN\triangleq\NN(\q_{NN};\wth): \R^{l}\times\R^{\Xi}\to\R^{n}$ represent the DNN, where $\q_{NN}\in\R^{l}$ is the DNN input vector, and $\wth\in\R^{\Xi}$ is the vector of trainable weights.
The architecture of $\NN(\q_{NN};\wth)$ will be defined in Section \ref{NN definition}.
According to the universal approximation theorem for DNNs \cite{RN33}, $\NN(\q_{NN};\wth)$ can approximate a nonlinear function $\mv g(\cdot)$ with an ideal weight vector $\wth^*$ on a compact subset $\Omega_{NN}\in\R^{l}$ to $\mv\epsilon$-accuracy, such that $\sup_{\q_{NN}\in\Omega_{NN}}\norm{ \NN(\q_{NN};\wth^*) - \mv g(\cdot)} = \mv\epsilon < \infty$.
Furthermore, the theorem states that the norm of $\wth^*$ is bounded, \ie $\norm{\wth^*}\le \bar\wth<\infty$.
In this study, $\wth^*$ is defined as a local optimal point, rather than a global optimal point.

In conclusion, the ideal control law $\mv\tau^*$ is expressed by the DNN approximation with the ideal weight vector $\NN^*\triangleq\NN(\q_{NN};\wth^*)$ as follows:
\begin{align}
    \mv\tau^*=& -(\NN^*+\mv\epsilon),
    \label{eq:control:ideal}
\end{align}
which is estimated online by
\begin{align}
    \mv\tau =& -\hat\NN,
    \label{eq:control:est}
\end{align}
where $\hat\NN\triangleq\NN(\q_{NN};\hat\wth)$, and  $\hat\wth$ is the estimated weight vector for $\wth^*$.

Substituting \eqref{eq:control:ideal} and \eqref{eq:control:est} into \eqref{eq:lya1:dot1}, the time derivative of $\mathcal V_1$ can be rewritten as
\begin{equation}
    \dot{\mathcal V}_1
    \le 
    -\lambda_\text{min}(\mm K)\norm{\fe}^2
    +
    \bar\tau_d\norm{\fe}
    +
    \fe^\top (\sat(\hat\NN)-\NN^*-\mv\epsilon)
    .
    \label{eq:lya1:dot2}
\end{equation}
Finally, one can conclude that the result of \eqref{eq:lya1:dot1} can be obtained by adapting $\hat\wth$ to $\wth^*$ (\ie $\hat\NN\to\NN^*$).

\subsection{Deep Neural Network (DNN) Model}\label{NN definition}

>>>>

\DeclareMathOperator{\myvec}{vec}
\DeclareMathOperator{\sat}{sat}
\DeclareMathOperator{\myrow}{row}

\newcommand*{\fe}{\mv r}

\newcommand*{\q}{\mv q}
\newcommand*{\dq}{\dot{\mv q}}
\newcommand*{\ddq}{\ddot{\mv q}}
\newcommand*{\qd}{\mv q_d}
\newcommand*{\dqd}{\dot{\mv q_d}}
\newcommand*{\ddqd}{\ddot{\mv q_d}}

\newcommand*{\NN}{\mv\Phi}
\newcommand*{\act}{\mv\phi}
\newcommand*{\wth}{\mv\theta}
\newcommand*{\wV}{\mm W}

\newcommand*{\estwth}{\widehat\wth}
\newcommand*{\estNN}{\widehat\NN}
\newcommand*{\estact}{\widehat\act}
\newcommand*{\estwV}{\widehat\wV}

\newcommand*{\idealwth}{\wth^*}
\newcommand*{\idealNN}{\NN^*}
\newcommand*{\idealact}{\act^*}
\newcommand*{\idealwV}{\wV^*}

@@

\section*{Notation}
In this study, the following notation is used:

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite[Definition 7.1.2]{Bernstein:2009aa}.
    \item $\mv x=[x_i]_{i\in[1,\cdots,n]}\in\R^n$ and $
        \mm A
        := 
        [a_{ij}]
        _{
            i\in[1,\cdots,n],j\in[1,\cdots ,]
        }\allowbreak\in\R^{n\times m}
        $ denotes a vector and a matrix.
    \item $\myrow_i(\mm A)$ denotes the $i\textsuperscript{th}$ row of the matrix $\mm A\in\R^{n\times m}$. 
    \item $\myvec(\mm A):= [\myrow_1(\mm A^\top)  ,\cdots,\myrow_m(\mm A^\top)  ]^\top   $ for $\mm A\in\R^{n\times m}$.
    % \item $\text{col}_i(A)$ and $\myrow_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\R^{n\times m}$, respectively.
    % \item $\myvec(A):= [\text{col}_1(A)^\top  ,\cdots,\text{col}_m(A)^\top  ]^\top   $ for $A\in\R^{n\times m}$.
    \item $\lambda_\text{min}(\mm A)$ denotes the minimum eigenvalue of the matrix $\mm A\in\R^{n\times n}$.
    \item $\mm I_n$ denotes the $n\times n$ identity matrix.
\end{itemize}

@@

\IEEEPARstart{M}{any} engineering systems, including those in aerospace, robotics, and automotive applications, can be modeled using Euler-Lagrange systems. 
These systems are governed by dynamic equations derived from energy principles and describe the motion of mechanical systems with constraints. 
In practice, however, such systems often exhibit uncertainties due to unmodeled dynamics, parameter variations, or external disturbances. 
These uncertainties can significantly degrade control performance and, in some cases, lead to instability. 
To address these challenges, adaptive control methods have been widely employed to ensure robust performance in the presence of system uncertainties \cite{Ioannou:2006aa, Tao:2003aa}.

More recently, neuro-adaptive control approaches have been introduced to approximate unknown system dynamics or entire control laws using neural networks (NNs) \cite{Farrell:2006aa}. 
NNs are well-known for their universal approximation property, which allows them to approximate any smooth function over a compact set with minimal error. 
Various types of NNs have been utilized in neuro-adaptive control, including simpler architectures like single-hidden layer (SHL) neural networks \cite{Ge:2010aa, Yesildirek:1995aa} and radial basis function (RBF) neural networks \cite{Liu:2013ab,Ge:2002aa}, as well as more complex models like deep neural networks (DNNs) \cite{Patil:2022aa} and their variations. 
SHL and RBF NNs are often employed to approximate uncertain system dynamics or controllers due to their simplicity \cite{Esfandiari:2014aa,Esfandiari:2015aa,Yesildirek:1995aa,Gao:2006aa}, while DNNs offer greater expressive power, making them more effective for complex system approximations \cite{Rolnick:2017aa}. 
Additionally, variations of DNNs, such as long short-term memory (LSTM) networks for time-varying dynamics \cite{Liu:2013ab} and physics-informed neural networks (PINNs) for leveraging physical system knowledge \cite{Hart:2024aa}, have further extended the capabilities of neuro-adaptive control systems.

@@

However, two significant challenges persist in using NNs for adaptive control. 
First, the boundedness of NN weights is not inherently guaranteed, which can result in unbounded outputs. 
When NN outputs are used directly in the control law, this may lead to excessive control inputs, violating input constraints. 
Such constraints are commonly encountered in industrial systems, where actuators are limited by physical and safety requirements in terms of amplitude, rate, or energy \cite{Esfandiari:2021aa}. 
Failing to address these constraints can degrade control performance or even destabilize the system.

Therefore, addressing these two key issues—ensuring weight boundedness and satisfying input constraints—is essential for the reliable design of neuro-adaptive controllers. 
he following section will provide a detailed review of the existing solutions to these challenges.

@@

A common challenge in neuro-adaptive control is maintaining the boundedness of the NN weights to ensure stability. 
In many studies, projection operators were employed to enforce upper bounds on the weight norms, ensuring that the weights do not grow unboundedly. 
For example, in \cite{Zhou:2023aa,Griffis:2023aa,Patil:2022aa}, projection operators were used to constrain the weight norms to remain below predetermined constants. 
However, these constants were often selected as large as possible due to the lack of theoretical guarantees regarding the global optimality of the weight values. 
While this approach ensured that the NN remained stable, it did not necessarily result in optimal performance.

In addition to projection operators, some studies utilized modification techniques like $\sigma$-modification \cite{Ge:2002aa} and $\epsilon$-modification \cite{Esfandiari:2015aa,Gao:2006aa}. 
These methods ensured that the NN weights remained within an invariant set by incorporating stabilizing functions into the adaptation law. 
Although these techniques were effective in ensuring boundedness and avoiding weight divergence, they similarly lacked a formal analysis of the optimality of the adapted weights, leaving room for improvement in terms of performance optimization.

@@

To address input saturation, many studies introduced auxiliary systems. 
These systems mitigated the effects of control input saturation by modifying the control strategy when saturation occurred. 
For instance, in \cite{Esfandiari:2014aa,Karason:1994aa,Esfandiari:2015aa}, auxiliary states were generated whenever input saturation was detected, and these states were incorporated into the adaptation law to adjust the NN weights accordingly. 
This approach helped the controller reduce input saturation by indirectly regulating the auxiliary states.

Alternatively, auxiliary states can also be used as feedback terms in the control law to directly compensate for the effects of input saturation constraints, as demonstrated in \cite{Arefinia:2020aa,He:2016aa,Peng:2020aa}. 
In \cite{Gao:2006aa}, the NN was used to approximate the desired control input, compensating for input saturation. 
However, these approaches typically handle input bound constraints on a per-input basis (\ie applying bounds to each scalar control variable individually), and may not account for more complex, nonlinear constraints, like input norm constraints, which are commonly found in physical systems such as robotic actuators or motor systems.

@@

To overcome these limitations, constrained optimization offers a promising approach. 
By formulating the neuro-adaptive control problem as an optimization problem with constraints, it is possible to adapt the NN weights while minimizing an objective function (\eg tracking error) subject to both weight norm and input constraints. 
Constrained optimization provides a theoretical framework for defining optimality and presents numerical methods for finding solutions that satisfy the constraints \cite{Nocedal:2006aa}.

In existing literature, constrained optimization techniques, such as the Augmented Lagrangian Method (ALM) \cite{Evens:2021aa} and the Alternating Direction Method of Multipliers (ADMM) \cite{Wang:2019aa,Taylor:2016aa}, have been used to train NNs offline. 
These methods impose constraints to address issues like gradient vanishing in backpropagation. 
However, to the best of the authors' knowledge, no prior work has applied constrained optimization to adaptive control systems with real-time weight adaptation. 
This gap suggests that constrained optimization could be key to addressing both weight norm boundedness and input constraints in a unified, theoretically grounded framework, particularly in real-time neuro-adaptive control.

@@

The main contributions of this study are listed as follows:
\begin{itemize}
    \item A constrained optimization-based neuro-adaptive controller (CoNAC) is developed using a DNN, where input constraints and the boundedness of NN weights are formulated as inequality constraints within the adaptation process.
    \item The weight adaptation laws in CoNAC are derived using constrained optimization theory to minimize the objective function while satisfying the inequality constraints. The adaptation laws ensure convergence of the weights to the first-order optimality conditions, specifically the Karush-Kuhn-Tucker (KKT) conditions.
    \item The forward sensitivity method is applied in CoNAC to accurately calculate the gradient of the objective function. \MSRY{WILL BE DELETED}
    %, enabling more precise real-time adaptation.
\end{itemize}

@@

The remainder of this paper is organized as follows. 
Section \ref{sec:Problem Formulation} presents the target system and control objective.
Section \ref{sec:ctrl design} introduces the proposed controller and the architecture of DNN in the controller. 
In Section \ref{sec:adap_laws}, the weight adaptation laws are developed.
Candidates of input constraints are presented in Section \ref{sec:appen:cstr}.
Section \ref{sec:stability} analyzes the stability of the proposed controller.
A comparative study of the four selected controllers, including the proposed controller, is reported in Section \ref{sec:sim}.
Finally, Section \ref{sec:conclusion} concludes the paper and discusses potential future work.
% The appendices provide the proof of the lemma used in the stability analysis.

%  SECTION PROBLEM FORMULATION ============================
\section{Problem Formulation}\label{sec:Problem Formulation}

@@

Consider an uncertain Euler-Lagrange system modeled as
\begin{equation}
    \mm M\ddq + \mm V_m + \mm F + \mm G + \mv\tau_d
    =
    \sat(\mv\tau)
    \label{eq:sys1}
\end{equation}
where $\q\in \R^n$ denotes the generalized coordinate, $\mv \tau_d$ represents disturbance and $\mv\tau\in\R^n$ denotes the control input. 
The terms $\mm M:=\mm M(\q)\in\R^{n\times n}$, $\mm V_m:=\mm V_m(\q,\dq)\in\R^{n\times n}$, $\mm F:=\mm F(\dq)\in\R^{n}$ and $\mm G:=\mm G(\q)$ represent the unknown inertia matrix, Coriolis/centripetal matrix, friction terms and gravity vector
The function $\sat(\cdot):\R^n\to\R^n$ represents the inherent physical limitations of the actuators and is defined as 
\begin{equation}
    \sat(\mv\tau)
    :=
    \begin{cases}
        \mv\tau, 
        & 
        \text{if } \Vert\mv\tau\Vert\le \bar\tau,
        \\
        \bar\tau
        \cdot
        \tfrac{\mv\tau}{\Vert\mv\tau\Vert}, 
        & 
        \text{otherwise},
    \end{cases}
    \label{eq:sat}
\end{equation}
where $\bar\tau$ denotes the maximum norm of control input.
To account for these limitations, it is essential to incorporate physically motivated constraints into the controller design.
Appendix \ref{sec:appen:cstr} introduces candidate constraints that can be applied to ensure compliance with these physical limitations.

@@

\begin{prop} \cite{Lewis:1998aa}
    The Coriolis/centripetal matrix $\mm V_m$ can always be selected so that the matrix $(\dot{\mm M}-2\mm V_m)$ is skew-symmetric.
    \label{prop:skew}
\end{prop}

\begin{prop}
    The disturbance $\mv\tau_d$ is bounded so that $\norm{\mv\tau_d}\le \bar\tau_d$.
    \label{prop:dis_bound}
\end{prop}

@@

The architecture of the proposed CoNAC is illustrated in Fig.~\ref{fig:ctrl:diagram}, consisting of: a DNN that functions as a neuro-adaptive controller, and a weight optimizer for the DNN. 
Section \ref{sec:sub:NAC} presents the neuro-adaptive controller along with the reference generator, and Section \ref{sec:sub:NN definition} defines the DNN model. The weight optimizer will be detailed in Section \ref{sec:sub:weight optimizer}.

\begin{figure*}[!t]
    \centering
    \includegraphics[width=0.7\linewidth]{fig/Controller.drawio.png}
    % \includesvg[width=0.75\linewidth]{Controller.drawio.svg}
    \caption{Architecture of the constrained optimization-based neuro-adaptive controller (CoNAC).}
    \label{fig:ctrl:diagram}
\end{figure*}

\subsection{Neuro-Adaptive Controller Design}\label{sec:sub:NAC}

First, let us define the filtered tracking error $\fe\in\R^n$ as 
\begin{equation}
    \fe := \dot{\mv e}-\mm\Lambda \mv e
    ,
    \label{eq:err:filtered}
\end{equation}
where $\mv e := \q - \qd$ denotes the tracking error, $\dot{\mv e} := \dq - \dqd$ represents the derivative of the tracking error, and $\mm\Lambda\in\R^{n\times n}$ is a user-designed positive definite matrix.
Since \eqref{eq:err:filtered} is stable system, it implies that $\mv e$ is bounded if $\fe$ is bounded.

Using $\fe$, the system dynamics \eqref{eq:sys1} can be rewritten as
\begin{equation}
    \mm M \dot\fe
    =
    -\mm V_m \fe
    -\mm K \fe
    + \mv f
    -\mv \tau_d + \sat(\mv\tau)
    ,
    \label{eq:sys:filtered}
\end{equation}
where $\mm K\in\R^{n\times n}$ denotes arbitrary symmetric positive definite matrix and $
    \mv f:= \mv f(\q,\dq,\dqd,\ddqd)
    =
    \mm K \fe
    +\mm M(-\ddqd+\mm\Lambda\dot{\mv e})
    +
    \mm V_m(-\dqd+\mm\Lambda\mv e)
    -
    \mm F
    -
    \mm G
    \in\R^n
$ denotes the lumped system uncertainty.

Consider the Lyapunov function $V_1:= \tfrac{1}{2}\fe^\top \mm M\fe$. 
Invoking Property \ref{prop:skew}, the time derivative of $V_1$ is
\begin{equation}
    \begin{aligned}
        \dot{V}_1
        =&
        \fe^\top \mm M (
            -\mm V_m \fe -\mm K \fe + \mv f
            -\mv \tau_d + \sat(\mv\tau)
        )
        +
        \tfrac{1}{2}
        \fe^\top \dot{\mm M}\fe
        \\
        =&
        -
        \fe^\top \mm K \fe 
        +
        \fe^\top (
            \mv f+\sat(\mv\tau)
            -\mv \tau_d
        )
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top(\dot{\mm M}- 2\mm V_m)\fe
        \\
        \le&
        -\lambda_\text{min}(\mm K)\norm{\fe}^2
        +
        \bar\tau_d\norm{\fe}
        +
        \fe^\top (\mv f+\sat(\mv \tau))
        \\
        \le&
        -\lambda_\text{min}(\mm K)\norm{\fe}^2
        +
        \bar\tau_d\norm{\fe}
        +
        \fe^\top (\sat(\mv\tau)-\mv\tau^*)
        ,
    \end{aligned}
    \label{eq:lya1:dot1}
\end{equation}
where $\mv\tau^*:= -\mv f$ is the ideal control input. 
Therefore, one can conclude that $\fe$ is exponentially stable so that $\lim_{t\to\infty}\norm{\fe}=\tfrac{\bar\tau_d}{\lambda_\text{min}(\mm K)}$, if $\mv \tau=-\mv\tau^*$ can be realized.
However, $\mv \tau^*$ is not available in practice, since $\mv f$ is unknown.

To overcome this issue, a DNN is employed to approximate $\mv\tau^*$.
Let $\NN:=\NN(\q_{NN};\wth): \R^{l}\times\R^{\Xi}\to\R^{n}$ represent the DNN, where $\q_{NN}\in\R^{l}$ is the DNN input vector, and $\wth\in\R^{\Xi}$ is the vector of trainable weights.
The architecture of $\NN(\q_{NN};\wth)$ will be defined in Section \ref{eq:DNN:architecture}.
According to the universal approximation theorem for DNNs \cite{Kidger:2020aa}, $\NN(\q_{NN};\wth)$ can approximate a nonlinear function $\mv g(\cdot)$ with an ideal weight vector $\idealwth$ on a compact subset $\Omega_{NN}\in\R^{l}$ to $\mv\epsilon$-accuracy, such that $\sup_{\q_{NN}\in\Omega_{NN}}\norm{ \NN(\q_{NN};\idealwth) - \mv g(\cdot)} = \mv\epsilon < \infty$.
Furthermore, the theorem states that the norm of $\idealwth$ is bounded, \ie $\norm{\idealwth}\le \bar\wth<\infty$.
In this study, $\idealwth$ is defined as a local optimal point, rather than a global optimal point.

In conclusion, the ideal control law $\mv\tau^*$ is expressed by the DNN approximation with the ideal weight vector $\idealNN:=\NN(\q_{NN};\idealwth)$ as follows:
\begin{align}
    \mv\tau^*=& \idealNN+\mv\epsilon,
    \label{eq:control:ideal}
\end{align}
which is estimated online by
\begin{align}
    \mv\tau =& \estNN,
    \label{eq:control:est}
\end{align}
where $\hat\NN:=\NN(\q_{NN};\estwth)$, and  $\estwth$ is the estimated weight vector for $\idealwth$.

Substituting \eqref{eq:control:ideal} and \eqref{eq:control:est} into \eqref{eq:lya1:dot1}, the time derivative of $\mathcal V_1$ can be rewritten as
\begin{equation}
    \dot{\mathcal V}_1
    \le 
    -\lambda_\text{min}(\mm K)\norm{\fe}^2
    +
    \bar\tau_d\norm{\fe}
    +
    \fe^\top (\sat(\hat\NN)-\idealNN-\mv\epsilon)
    .
    \label{eq:lya1:dot2}
\end{equation}
Finally, one can conclude that the result of \eqref{eq:lya1:dot1} can be obtained by adapting $\estwth$ to $\idealwth$ (\ie $\hat\NN\to\idealNN$).

\subsection{Deep Neural Network (DNN) Model}\label{sec:sub:NN definition}
